Lineage Analysis of CRISPR and shRNA Screens

### 2018-10-16 lineages_18Q3.Rmd
update rmd

```{r}
makeCRISPRgrob <- function(g) {

  gene <- filter(crispr_data, Hugo_Symbol == g)

  # Mutation status
  crispr_color <- as.character(gene$Color)
  names(crispr_color) <- gene$Mutation_Status
  sig <- filter(crispr_signif, Hugo_Symbol == g)
  plot_mut <- ggplot(data = gene, aes(x = Mutation_Status, y = Score, color = Mutation_Status)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(alpha = 0.3, size = 0.7, position = position_jitter(w = 0.05)) +
    scale_color_manual(values = crispr_color) +
    geom_hline(yintercept = 0, linetype = 2, lwd = 0.3) +
    theme_light() +
    theme(legend.position = "none") +
    labs(x = paste0(g, " Mutation Status"), y = "CERES Score",
         title = "Mutation Status",
         subtitle = paste0("Wilcoxon test:\n- BH-corrected p-value: ", sig$p.adj.short, "\n- Uncorrected p-value: ", sig$p.short, "\nMutant lines harbor deleterious mutations."))

  # Copy number
  lo_cn <- range(gene$Copy_Number[!is.na(gene$Copy_Number)])[1]
  hi_cn <- range(gene$Copy_Number[!is.na(gene$Copy_Number)])[2]
  mid_cn <- (hi_cn - lo_cn) / 2
  plot_cn <- ggplot(data = gene, aes(x = Copy_Number, y = Score, color = Mutation_Status)) +
    geom_point(size = 0.5, alpha = 0.5) +
    geom_smooth(method = "lm", size = 0.5) +
    geom_hline(yintercept = 0, linetype = 2, lwd = 0.3) +
    scale_color_manual(values = crispr_color) +
    stat_cor(method = "pearson", show.legend = FALSE, label.x = c((lo_cn + mid_cn) / 2, (hi_cn - mid_cn) / 2 + mid_cn), label.y = max(gene$Score)) +
    theme(legend.position = "none") +
    labs(y = "CERES Score", color = "Mutation Status",
         x = paste0(g, " Theoretical Copy number"), title = "Copy Number",
         subtitle = "r: Pearson correlation coeffcient")

  # Gene expression
  lo_ge <- range(gene$RPKM_log2[!is.na(gene$RPKM_log2)])[1]
  hi_ge <- range(gene$RPKM_log2[!is.na(gene$RPKM_log2)])[2]
  mid_ge <- (hi_ge - lo_ge) / 2
  plot_ge <- ggplot(data = gene, aes(x = RPKM_log2, y = Score, color = Mutation_Status)) +
    geom_point(size = 0.5, alpha = 0.5) +
    geom_smooth(method = "lm", size = 0.5) +
    geom_hline(yintercept = 0, linetype = 2, lwd = 0.3) +
    scale_color_manual(values = crispr_color) +
    stat_cor(method = "pearson", show.legend = FALSE, label.x = c((lo_ge + mid_ge) / 2, (hi_ge - mid_ge) / 2 + mid_ge), label.y = max(gene$Score)) +
    theme(legend.position = "none") +
    labs(y = "CERES Score", color = "Mutation Status",
         x = paste0(g, " Gene Expression [log2(RPKM)]"),
         title = "Gene Expression",
         subtitle = "r: Pearson correlation coeffcient")

  # Cell line lineage
  plot_tissue <- ggplot(data = gene, aes(x = primary_tissue, y = Score, color = Mutation_Status)) +
    geom_point(alpha = 0.5) +
    scale_color_manual(values = crispr_color) +
    geom_hline(yintercept = 0, linetype = 2, lwd = 0.3) +
    scale_y_continuous(sec.axis = sec_axis(~ .)) +
    coord_flip() +
    theme(legend.position = "none") +
    labs(y = "CERES Score", x = "Primary Tissue",
         color = "Mutation Status", title = g)

@@

## Make tissue plot

```{r}
makeCRISPRtissueplot <- function(g) {
  sig <- filter(crispr_signif_tissue, Hugo_Symbol == g)
  if(nrow(sig) == 0) {
    return(NULL)
  }
  else {
    gene <- filter(crispr_data, Hugo_Symbol == g)
    gene <- merge(gene, sig, by = c("Hugo_Symbol", "primary_tissue"))
    gene <- mutate(gene, primary_tissue = reorder(primary_tissue, p, mean))
    crispr_color <- as.character(gene$Color)
    names(crispr_color) <- gene$Mutation_Status
    score_range <- abs(range(gene$Score)[2] - range(gene$Score)[1])
    round_accuracy <- ifelse(score_range <= 2, 0.25,
                             ifelse(score_range <= 3, 0.5, 1.0))

    plot <- ggplot(data = gene, aes(x = primary_tissue, y = Score)) +
      geom_point(alpha = 0.5, mapping = aes(color = Mutation_Status)) +
      scale_color_manual(values = crispr_color) +
      coord_cartesian(y = c(min(gene$Score), round_any(x = max(gene$Score), accuracy = round_accuracy, f = ceiling))) +
      geom_signif(data = gene, mapping = aes(xmin = primary_tissue, xmax = primary_tissue, annotations = paste(p.short, "\n", p.adj.short), y_position = max(gene$Score)), manual = TRUE, tip_length   = 0, size = 0, textsize = 3) +
      theme(legend.position = "top", axis.text.x = element_text(angle = 70, hjust = 1, size = 10)) +
      labs(y = "CERES Score",
           x = "Primary Tissue",
           color = "Mutation Status",
           title = paste0(g, ": CERES score by cell line primary tissue"),
           subtitle = "Primary tissues sorted by increasing p-value; labels indicate unadjusted p-value / BH-corrected p-value.\nMutant lines harbor deleterious mutations.")
    return(plot)
  }
}
```

@@

@@

### DepMap cell line metadata

@@

# Merge all data
drug_data <- merge(merge(ccle_melt, ctrp_melt, by = c("accession_id", "CID", "Drug"), all = TRUE), gdsc_melt, by = c("accession_id", "CID", "Drug"), all = TRUE)
drug_data <- merge(drug_data, ccl_converter, by = "accession_id", all.x = TRUE)
drug_data <- merge(drug_data, maf_df, by = c("CCLE_Name", "Broad_ID"))
drug_data$Hugo_Symbol <- factor(drug_muts$Hugo_Symbol)

saveRDS(drug_data, "./../crispr_lineages_giant_files/drug_data_18Q3.rds", compress = "xz")
```

@@

### Grouped by gene and lineage: Non-silent mutant vs. Other

@@

lvlA_signif_genes_plot <- ggplot(data = lvlA_genes_signif, aes(x = Mutation_Status_Nonsilent, y = Score)) +
  facet_wrap(~ Hugo_Symbol, nrow = 1, scales = "free_y") +
  geom_jitter(aes(color = Mutation_Status_Nonsilent), alpha = 0.5, width = 0.3) +
  geom_boxplot(color = "black", fill = NA, outlier.shape = NA) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  scale_color_manual(values = lvlA_genes_signif_color) + 
  geom_text(data = lvlA_genes_text, mapping = aes(x = 1.5, y = Text_y, label = p.signif.adj), nudge_y = 0.2) +
  theme(legend.position = "top", axis.text.x = element_text(angle = 65, hjust = 1, vjust = 1, size = 10)) +
  labs(y = "CERES Score",
       x = "Lineage",
       color = "Mutation Status",
       title = "CERES scores for significant Level A G2P genes",
       subtitle = "BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001\nGenes are sorted by increasing p-value.")
lvlA_signif_genes_plot

@@

```{r}
test <- '{"feature_names": "ALK positive", "tags": "test", "genes": "ALK"}'

fromJSON(test)
```

@@

```{r fig.height=12, fig.width=12, include=FALSE, eval=FALSE}
makeCRISPRgrob("KRAS")
```

Plot code:
```{r, eval=FALSE}
crispr_data_cgc <- filter(crispr_data, Hugo_Symbol %in% cgc$Hugo_Symbol)
crispr_all_genes <- unique(crispr_data_cgc$Hugo_Symbol)
crispr_all_groups <- lapply(crispr_all_genes, makeCRISPRgrob)
crispr_all_groups_paths <- paste0(crispr_all_genes, "_crispr_grouped.png")
pwalk(list(crispr_all_groups_paths, crispr_all_groups), ggsave, path = "./plots_18Q3/crispr_grouped_plots_cgc", dpi = 300, width = 12, height = 12, units = "in")
```

@@

## Sorted tissue plots

```{r fig.height=12, fig.width=12, include=FALSE, eval=FALSE}
makeCRISPRtissueplot("KRAS")
```

Plot code:
```{r, eval=FALSE}
crispr_data_cgc <- filter(crispr_data, Hugo_Symbol %in% cgc$Hugo_Symbol)
crispr_all_genes <- unique(crispr_data_cgc$Hugo_Symbol)
crispr_all_lins <- lapply(crispr_all_genes, makeCRISPRtissueplot)
crispr_all_lins_paths <- paste0(crispr_all_genes, "_crispr_tissue.png")
pwalk(list(crispr_all_lins_paths, crispr_all_lins), ggsave, path = "./plots_18Q3/crispr_tissue_plots_cgc", dpi = 300, width = 12, height = 12, units = "in")
```

Data management:
```{r, eval=FALSE}
crispr_all_bygene <- paste0(list.files("./plots_18Q3/crispr_tissue_plots_cgc", full.names = TRUE))
names(crispr_all_bygene) <- str_replace_all(crispr_all_bygene, c("_crispr_tissue.png" = "", "./plots_18Q3/crispr_tissue_plots_cgc/" = ""))
crispr_bygene_order <- c(intersect(crispr_signif$Hugo_Symbol, names(crispr_all_bygene)), setdiff(names(crispr_all_bygene), crispr_signif$Hugo_Symbol))
crispr_all_bygene <- crispr_all_bygene[match(crispr_bygene_order, names(crispr_all_bygene))]
bsselect(crispr_all_bygene, type = "img", selected = "KRAS", live_search = TRUE, show_tick = TRUE, height = 300, frame_height = 275)
```

>>>>

```{r}
makeCRISPRgrob <- function(g, filt_method) {
  if(filt_method == "lvlA") {
    gene <- filter(crispr_data_lvlA, Hugo_Symbol == g)
    sig <- filter(crispr_signif_lvlA, Hugo_Symbol == g)
  }
  if(filt_method == "all") {
    gene <- filter(crispr_data_ptmuts, Hugo_Symbol == g)
    sig <- filter(crispr_signif_nonsilent, Hugo_Symbol == g)
  }

  # Mutation status
  crispr_color <- as.character(gene$Color_Nonsilent)
  names(crispr_color) <- gene$Mutation_Status_Nonsilent

  plot_mut <- ggplot(data = gene, aes(x = Mutation_Status_Nonsilent, y = Score, color = Mutation_Status_Nonsilent)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(alpha = 0.3, size = 0.7, position = position_jitter(w = 0.05)) +
    scale_color_manual(values = crispr_color) +
    geom_hline(yintercept = 0, linetype = 2, lwd = 0.3) +
    theme_light() +
    theme(legend.position = "none") +
    labs(x = paste0(g, " Mutation Status"), y = "CERES Score",
         title = "Mutation Status",
         subtitle = paste0("Wilcoxon test:\n- BH-corrected p-value: ", sig$p.adj.short, "\n- Uncorrected p-value: ", sig$p.short))

  # Copy number
  lo_cn <- range(gene$Copy_Number[!is.na(gene$Copy_Number)])[1]
  hi_cn <- range(gene$Copy_Number[!is.na(gene$Copy_Number)])[2]
  mid_cn <- (hi_cn - lo_cn) / 2
  plot_cn <- ggplot(data = gene, aes(x = Copy_Number, y = Score, color = Mutation_Status_Nonsilent)) +
    geom_point(size = 0.5, alpha = 0.5) +
    geom_smooth(method = "lm", size = 0.5) +
    geom_hline(yintercept = 0, linetype = 2, lwd = 0.3) +
    scale_color_manual(values = crispr_color) +
    stat_cor(method = "pearson", show.legend = FALSE, label.x = c((lo_cn + mid_cn) / 2, (hi_cn - mid_cn) / 2 + mid_cn), label.y = max(gene$Score)) +
    theme(legend.position = "none") +
    labs(y = "CERES Score", color = "Mutation Status",
         x = paste0(g, " Theoretical Copy number"), title = "Copy Number",
         subtitle = "r: Pearson correlation coeffcient")

  # Gene expression
  lo_ge <- range(gene$RPKM_log2[!is.na(gene$RPKM_log2)])[1]
  hi_ge <- range(gene$RPKM_log2[!is.na(gene$RPKM_log2)])[2]
  mid_ge <- (hi_ge - lo_ge) / 2
  plot_ge <- ggplot(data = gene, aes(x = RPKM_log2, y = Score, color = Mutation_Status_Nonsilent)) +
    geom_point(size = 0.5, alpha = 0.5) +
    geom_smooth(method = "lm", size = 0.5) +
    geom_hline(yintercept = 0, linetype = 2, lwd = 0.3) +
    scale_color_manual(values = crispr_color) +
    stat_cor(method = "pearson", show.legend = FALSE, label.x = c((lo_ge + mid_ge) / 2, (hi_ge - mid_ge) / 2 + mid_ge), label.y = max(gene$Score)) +
    theme(legend.position = "none") +
    labs(y = "CERES Score", color = "Mutation Status",
         x = paste0(g, " Gene Expression [log2(RPKM)]"),
         title = "Gene Expression",
         subtitle = "r: Pearson correlation coeffcient")

  # Cell line lineage
  plot_tissue <- ggplot(data = gene, aes(x = primary_tissue, y = Score, color = Mutation_Status_Nonsilent)) +
    geom_point(alpha = 0.5) +
    scale_color_manual(values = crispr_color) +
    geom_hline(yintercept = 0, linetype = 2, lwd = 0.3) +
    scale_y_continuous(sec.axis = sec_axis(~ .)) +
    coord_flip() +
    theme(legend.position = "none") +
    labs(y = "CERES Score", x = "Primary Tissue",
         color = "Mutation Status", title = g)

@@

@@

## DepMap cell line metadata

@@

@@

# Merge all data
drug_data <- merge(merge(ccle_melt, ctrp_melt, by = c("accession_id", "CID", "Drug"), all = TRUE), gdsc_melt, by = c("accession_id", "CID", "Drug"), all = TRUE)
drug_data <- merge(drug_data, ccl_converter, by = "accession_id", all.x = TRUE)
drug_data <- merge(drug_data, maf_df, by = c("CCLE_Name", "Broad_ID"))
drug_data$Hugo_Symbol <- factor(drug_data$Hugo_Symbol)

saveRDS(drug_data, "./data_munging/drug_data_18Q3.rds", compress = "xz")
```

```{r}
drug_data <- readRDS("./data_munging/drug_data_18Q3.rds")
```

@@

@@

lvlA_signif_genes_plot <- ggplot(data = lvlA_genes_signif, aes(x = Mutation_Status_Nonsilent, y = Score)) +
  facet_wrap(~ Hugo_Symbol, nrow = 1, scales = "free_y") +
  geom_jitter(aes(color = Mutation_Status_Nonsilent), alpha = 0.5, width = 0.3) +
  geom_boxplot(color = "black", fill = NA, outlier.shape = NA) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  scale_color_manual(values = lvlA_genes_signif_color) + 
  geom_text(data = lvlA_genes_text, mapping = aes(x = 1.5, y = Text_y, label = p.signif.adj), nudge_y = 0.2) +
  theme(legend.position = "top") +
  labs(y = "CERES Score",
       x = "Lineage",
       color = "Mutation Status",
       title = "CERES scores for significant Level A G2P genes",
       subtitle = "BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001\nGenes are sorted by increasing p-value.")
lvlA_signif_genes_plot

@@

## Level A: Per-gene grouped plots

Plot code:
```{r, eval=FALSE}
crispr_lvlA_groups <- lapply(as.character(genes_lvlA), makeCRISPRgrob, filt_method = "lvlA")
crispr_lvlA_groups_paths <- paste0(genes_lvlA, "_crispr_grouped.png")
pwalk(list(crispr_lvlA_groups_paths, crispr_lvlA_groups), ggsave, path = "./plots_18Q3/crispr_grouped_plots_lvlA", dpi = 300, width = 12, height = 12, units = "in")
```

Data management:
```{r, eval=TRUE}
crispr_lvlA_bygene <- paste0(list.files("./plots_18Q3/crispr_grouped_plots_lvlA", full.names = TRUE))
names(crispr_lvlA_bygene) <- str_replace_all(crispr_lvlA_bygene, c("_crispr_grouped.png" = "", "./plots_18Q3/crispr_grouped_plots_lvlA/" = ""))
crispr_lvlA_order <- c(intersect(crispr_signif_lvlA$Hugo_Symbol, names(crispr_lvlA_bygene)), setdiff(names(crispr_lvlA_bygene), crispr_signif_lvlA$Hugo_Symbol))
crispr_lvlA_bygene <- crispr_lvlA_bygene[match(crispr_lvlA_order, names(crispr_lvlA_bygene))]
bsselect(crispr_lvlA_bygene, type = "img", selected = "KRAS", live_search = TRUE, show_tick = TRUE, height = 300, frame_height = 275)
```

## All genes: Per-gene grouped plots

Plot code:
```{r, eval=FALSE}
crispr_all_groups <- lapply(unique(crispr_data_ptmuts$Hugo_Symbol), makeCRISPRgrob, filt_method = "all")
crispr_all_groups_paths <- paste0(unique(crispr_data_ptmuts$Hugo_Symbol), "_crispr_grouped.png")
pwalk(list(crispr_lvlA_groups_paths, crispr_lvlA_groups), ggsave, path = "./plots_18Q3/crispr_grouped_plots", dpi = 300, width = 12, height = 12, units = "in")
```

Data management:
```{r, eval=TRUE}
crispr_all_bygene <- paste0(list.files("./plots_18Q3/crispr_grouped_plots", full.names = TRUE))
names(crispr_all_bygene) <- str_replace_all(crispr_all_bygene, c("_crispr_grouped.png" = "", "./plots_18Q3/crispr_grouped_plots_lvlA/" = ""))
crispr_all_order <- c(intersect(crispr_signif_nonsilent$Hugo_Symbol, names(crispr_all_bygene)), setdiff(names(crispr_all_bygene), crispr_signif_nonsilent$Hugo_Symbol))
crispr_all_bygene <- crispr_all_bygene[match(crispr_all_order, names(crispr_all_bygene))]
bsselect(crispr_all_bygene, type = "img", selected = "KRAS", live_search = TRUE, show_tick = TRUE, height = 300, frame_height = 275)
```

## JSON

@@

```{r fig.height=12, fig.width=12, include=FALSE, eval=FALSE}
makeCRISPRgrob("KRAS", filt_method = "all")
```

Plot code:
```{r, eval=FALSE}
crispr_lvlA_genes <- unique(crispr_data_cgc$Hugo_Symbol)
crispr_all_groups <- lapply(crispr_all_genes, makeCRISPRgrob)
crispr_all_groups_paths <- paste0(crispr_all_genes, "_crispr_grouped.png")
pwalk(list(crispr_all_groups_paths, crispr_all_groups), ggsave, path = "./plots_18Q3/crispr_grouped_plots_cgc", dpi = 300, width = 12, height = 12, units = "in")
```

@@
